Mass surveillance in cyberspace and the lost art of keeping a secret\thanks{The ideas discussed here first appeared as a DCSS project conference talk and were subsequently fleshed out in paper due to serendipitous encounters fostered by twitter.}

### 2015-07-08 edemocracy.tex
Final commit
Minor typos spotted.

\title{Mass surveillance in cyberspace and the lost art of keeping a secret\thanks{The ideas presented in this paper first appeared as a talk at the 2015 Conference on Digital Citizenship and Surveillance Society in Cardiff in June 2015 and subsequently refined and developed into this paper via serendipitous encounters that were enabled by the social medium of Twitter. The work has been sponsored in part by the European Commission through project NIFTy (ISEC 2012 Action Grants HOME/2012/ISEC/INT/4000003892).}}
\subtitle{Policy lessons for nation-states after the Snowden leaks}

@@

We referred earlier to CCTV as an example of a surveillance technology that has been relatively successful, from a technology acceptance and use point of view. That is particularly apparent in the UK, where it is estimated that there were over 4 million cameras in operation, although police investigations suggest that this may be an over-estimation~\cite{HomeOffice}. It is interesting to note that CCTV overall is really a collection of ad hoc installations of security cameras that offer access to footage of varying quality. These include public spaces monitoring as well as private surveillance systems for both corporate and personal property.

@@

The simplicity of purpose of a video surveillance system and the ability to relate its operation to a societal challenge (in this case football violence and public safety in grounds) established the technology in the collective social mind as something intrusive, yet necessary. This facilitated take up across public spaces of local councils (e.g. car parks), motorways, even on board means of transport, as well as in private business and buildings. Market forces and accessibility to the technology enabled everyone that may have had interest in it to install and operate such systems, creating space for the general public to serve as co-creators of the overall CCTV system-of-systems.

>>>>

\title{Mass surveillance in cyberspace and the lost art of keeping a secret\thanks{The ideas presented in this paper first appeared as a talk at the 2015 Conference on Digital Citizenship and Surveillance Society in Cardiff in June 2015 and subsequently refined and developed into this paper via serendipitous encounters that were enabled by the social medium of Twitter. The work has been sponsored in part by the European Commission through project NIFTy (ISEC 2012 Action Grants HOME/2012/ISEC/INT/4000003892).}}
\subtitle{Policy lessons for Government after the Snowden leaks}

@@

We referred earlier to CCTV as an example of a surveillance technology that has been relatively successful, from a technology acceptance and use point of view. That is particularly apparent in the UK, where it is estimated that there are over 4 million cameras in operation, although police investigations suggest that this may be an over-estimation~\cite{HomeOffice}. It is interesting to note that CCTV overall is really a collection of ad hoc installations of security cameras that offer access to footage of varying quality. These include public spaces monitoring as well as private surveillance systems for both corporate and personal property.

@@

The simplicity of purpose of a video surveillance system and the ability to relate its operation to a societal challenge (in this case football violence and public safety in grounds) established the technology in the collective social mind as something intrusive, yet necessary. This facilitated take up across public spaces of local councils (e.g. car parks), motorways, even on board means of transport, as well as in private businesses and buildings. Market forces and accessibility to the technology enabled everyone that may have had interest in it to install and operate such systems, creating space for the general public to serve as co-creators of the overall CCTV system-of-systems.
